---
title: "🌸 Cherry Blossom Climate Phenology & Predictive Forecasting Hub"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(forecast)
library(urca)
library(ggfortify)
library(shiny)

# =========================================================================
# 1. CORE DATA PIPELINE INGESTION (004 & 005)
# =========================================================================
bloom_raw <- read.csv("005.csv", header = FALSE, fileEncoding = "CP932", stringsAsFactors = FALSE, na.strings = c("", "NA"))
b_years   <- as.numeric(as.character(bloom_raw[2, grep("^19|^20", as.character(bloom_raw[2, ]))]))
b_idx     <- grep("^19|^20", as.character(bloom_raw[2, ]))

b_list <- list()
for (i in 3:nrow(bloom_raw)) {
  s_id <- as.character(bloom_raw[i, 1])
  if (!is.na(s_id) && grepl("^[0-9]+", s_id)) {
    for (y in 1:length(b_years)) {
      val <- trimws(as.character(bloom_raw[i, b_idx[y]]))
      if (!is.na(val) && val != "" && val != "0" && grepl("^[0-9]+", val)) {
        v <- as.numeric(val); m <- floor(v/100); d <- v%%100
        if (m >= 1 && m <= 12 && d >= 1 && d <= 31) {
          doy <- as.numeric(format(as.Date(paste(b_years[y], m, d, sep="-")), "%j"))
          b_list[[length(b_list)+1]] <- data.frame(Station_ID=s_id, Year=b_years[y], Bloom_DOY=doy, stringsAsFactors = FALSE)
        }
      }
    }
  }
}
df_bloom <- do.call(rbind, b_list)

flow_raw <- read.csv("004.csv", header = FALSE, fileEncoding = "CP932", stringsAsFactors = FALSE, na.strings = c("", "NA"))
f_years  <- as.numeric(as.character(flow_raw[2, grep("^19|^20", as.character(flow_raw[2, ]))]))
f_idx    <- grep("^19|^20", as.character(flow_raw[2, ]))

f_list <- list()
for (i in 3:nrow(flow_raw)) {
  s_id <- as.character(flow_raw[i, 1])
  if (!is.na(s_id) && grepl("^[0-9]+", s_id)) {
    for (y in 1:length(f_years)) {
      val <- trimws(as.character(flow_raw[i, f_idx[y]]))
      if (!is.na(val) && val != "" && val != "0" && grepl("^[0-9]+", val)) {
        v <- as.numeric(val); m <- floor(v/100); d <- v%%100
        if (m >= 1 && m <= 12 && d >= 1 && d <= 31) {
          doy <- as.numeric(format(as.Date(paste(f_years[y], m, d, sep="-")), "%j"))
          f_list[[length(f_list)+1]] <- data.frame(Station_ID=s_id, Year=f_years[y], Flow_DOY=doy, stringsAsFactors = FALSE)
        }
      }
    }
  }
}
df_flow <- do.call(rbind, f_list)

calc_slopes <- function(df, val_col) {
  s_list <- unique(df[["Station_ID"]])
  out <- list()
  for (s in s_list) {
    sub <- df[df[["Station_ID"]] == s, ]
    if (nrow(sub) >= 15) {
      slope <- cov(sub[["Year"]], sub[[val_col]]) / var(sub[["Year"]])
      out[[s]] <- data.frame(Station_ID=s, Rate=slope, stringsAsFactors = FALSE)
    }
  }
  return(do.call(rbind, out))
}

bloom_slopes <- calc_slopes(df_bloom, "Bloom_DOY")
flow_slopes  <- calc_slopes(df_flow, "Flow_DOY")

comp <- inner_join(bloom_slopes, flow_slopes, by="Station_ID", suffix=c("_Bloom", "_Flow")) %>%
  mutate(Total_Bloom_Shift = Rate_Bloom * 73)

city_lookup <- data.frame(
  Station_ID = c("626", "401", "406", "407", "409", "412", "413", "417", "418", "420", "423", "428", "430", "433", "435", "575", "581", "582", "584", "585", "587", "588", "590", "595", "759", "762", "827", "891"),
  City_Name  = c("Tokyo (Central)", "Wakkanai", "Rumoi", "Asahikawa", "Abashiri", "Sapporo", "Nemuro", "Hakodate", "Kushiro", "Muroran", "Urakawa", "Esashi", "Hakodate-Base", "Kutchan", "Mombetsu", "Fukui", "Tsuruga", "Nagoya", "Gifu", "Takayama", "Shizuoka", "Omaezaki", "Mishima", "Kofu", "Kyoto", "Osaka", "Nagasaki", "Kagoshima"),
  stringsAsFactors = FALSE
)

menu_choices        <- city_lookup[["Station_ID"]]
names(menu_choices) <- city_lookup[["City_Name"]]
```
# Historical Trends & Spatial Matrices {data-icon="fa-chart-area" data-orientation=rows}

## Row {data-height=110}

### Network-Wide Peak Bloom Shift Rate

```{r}
valueBox(
  value = paste(round(mean(comp[["Rate_Bloom"]]), 4), "Days/Year"),
  icon = "fa-clock",
  color = "#d81b60"
)
```

### Cumulative Peak Advancement (73 Years)

```{r}
valueBox(
  value = paste(round(abs(mean(comp[["Total_Bloom_Shift"]])), 1), "Days Earlier"),
  icon = "fa-calendar-check",
  color = "#fad0c4"
)
```

### Initial Flowering Shift Velocity

```{r}
valueBox(
  value = paste(round(mean(comp[["Rate_Flow"]]), 4), "Days/Year"),
  icon = "fa-seedling",
  color = "#4db6ac"
)
```

## Row {data-height=400}

### Peak Bloom Macro Trajectory Tracking

```{r}
target_stations <- unique(df_bloom[["Station_ID"]])[1:4]
plot_subset     <- df_bloom %>% filter(Station_ID %in% target_stations)

ggplot(plot_subset, aes(x = Year, y = Bloom_DOY, color = as.factor(Station_ID))) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, linewidth = 1.2) +
  scale_color_brewer(palette = "Set1") +
  labs(
    subtitle = expression(paste("Global Matrix: ", italic(Y) == beta + beta*italic(X) + beta*italic(X)^2 + epsilon)),
    x = "Observation Year", y = "Julian Day of Year", color = "Station ID"
  ) +
  theme_minimal()
```

### Regional Response Contrast Dispersion Proxy

```{r}
map_data <- comp %>%
  mutate(
    Station_Num = as.numeric(gsub("[^0-9]", "", as.character(Station_ID))),
    Lat_Proxy   = ifelse(Station_Num < 420, 43.0 + (Station_Num %% 5)*0.2, 
                         ifelse(Station_Num < 500, 38.2 + (Station_Num %% 5)*0.2, 35.6 - (Station_Num %% 10)*0.3)),
    Lon_Proxy   = ifelse(Station_Num < 420, 141.3 + (Station_Num %% 5)*0.2, 
                         ifelse(Station_Num < 500, 140.8 + (Station_Num %% 5)*0.2, 139.6 - (Station_Num %% 10)*0.3))
  )

ggplot(map_data, aes(x = Lon_Proxy, y = Lat_Proxy, fill = Total_Bloom_Shift)) +
  geom_point(size = 4.5, shape = 21, color = "#2c3e50", stroke = 0.7, alpha = 0.9) +
  scale_fill_gradientn(colors = c("#d81b60", "#f48fb1", "#e0e0e0", "#4db6ac", "#00695c"), name = "Shift Days") +
  labs(subtitle = "Latitudinal Matrix Shift Velocity Scatter Array", x = "Proxy Longitude Vector", y = "Proxy Latitude Vector") +
  theme_minimal() + 
  theme(panel.background = element_rect(fill = "#fdfbfb", color = NA), panel.grid.minor = element_blank())
```

## Row {data-height=300}

### Searchable Station Inventory Metrics

```{r}
final_table_data <- comp %>%
  left_join(city_lookup, by = "Station_ID") %>%
  mutate(City_Name = ifelse(is.na(City_Name), paste("Regional Station", Station_ID), City_Name)) %>%
  select(ID = Station_ID, Location = City_Name, Rate = Rate_Bloom, `Total Shift` = Total_Bloom_Shift) %>%
  mutate(Rate = round(Rate, 4), `Total Shift` = round(`Total Shift`, 1)) %>%
  arrange(`Total Shift`)

base_table <- datatable(final_table_data, filter = 'top', options = list(pageLength = 5, autoWidth = TRUE, columnDefs = list(list(className = 'dt-center', targets = 0:3))), rownames = FALSE)

formatStyle(table = base_table, columns = 'Total Shift', backgroundColor = styleInterval(c(-10, -5), c('#fbc2eb', '#fad0c4', '#fdfbfb')), color = '#2c3e50')
```


# Predictive Stochastic Forecasting {data-icon="fa-magic" data-orientation=columns}

## Sidebar {.sidebar}

### Dynamic Selection Terminal

```{r}
selectInput(
  inputId  = "selected_station",
  label    = "Choose Target Meteorological Station:",
  choices  = menu_choices,
  selected = "626"
)
```

Use this panel to dynamically force the background computing environment to isolate a specific historical station chronology timeline. 

Selecting an ID triggers an automated recalculation matrix:
* Recopies local data limits
* Re-evaluates unit root presence
* Identifies optimal (p, d, q) lags
* Projects confidence horizons out to 2036

## Column {data-width=550}

### Automated Stochastic ARIMA Projections (10-Year Boundary View)

```{r}
renderPlot({
  station_data <- df_bloom %>% 
    filter(Station_ID == input[["selected_station"]]) %>% 
    arrange(Year)
  
  station_ts <- ts(station_data[["Bloom_DOY"]], start = min(station_data[["Year"]]), frequency = 1)
  
  fit_arima <- auto.arima(station_ts, seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
  fc_array  <- forecast(fit_arima, h = 10)
  
  # Gently change the color mapping without breaking autoplot's internal structure
  autoplot(fc_array) +
    scale_colour_discrete(h = c(330, 350), c = 80, l = 50) + 
    scale_fill_discrete(h = c(340, 360), c = 60, l = 85) +
    labs(
      subtitle = paste("Optimal Model Configuration:", fit_arima$method, "| Shading: 80% & 95% Confidence Horizons"),
      x = "Observation Timeline", y = "Julian Day of Year"
    ) +
    theme_minimal()
})
```


## Column {data-width=450}

### Dynamic Time-Series Econometric Interpretation

```{r}
renderUI({
  station_data <- df_bloom %>% 
    filter(Station_ID == input[["selected_station"]]) %>% 
    arrange(Year)
  
  station_ts <- ts(station_data[["Bloom_DOY"]], start = min(station_data[["Year"]]), frequency = 1)
  
  fit_arima <- auto.arima(station_ts, seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
  adf_test  <- ur.df(station_ts, type = "trend", lags = 1)
  
  loc_name  <- city_lookup %>% filter(Station_ID == input[["selected_station"]]) %>% pull(City_Name)
  if(length(loc_name) == 0) loc_name <- paste("Station", input[["selected_station"]])
  
  HTML(paste0(
    "<strong>Target Focus Location: </strong>", loc_name, " (ID: ", input[["selected_station"]], ")<br><br>",
    "<strong>The Unit Root Diagnostic (Augmented Dickey-Fuller Specification)</strong><br>",
    "The historical timeline for this location yields a calculated ADF test statistic of <strong>", 
    round(summary(adf_test)@teststat, 4), "</strong> against a standard 5% Critical Value threshold of <strong>-3.45</strong>. If the test statistic is more negative than -3.45, we reject the null hypothesis of a unit root, confirming the series is trend-stationary.<br><br>",
    "<strong>Model Selection & Coefficients: </strong><br>",
    "The underlying Akaike Information Criterion system selected an optimal <strong>", fit_arima$method, "</strong> framework topology profile matrix for this dataset.<br><br>",
    "<strong>Deterministic vs. Stochastic Modeling Dynamics</strong><br>",
    "Unlike a rigid linear trend that projects an infinite downward slide, the automated stochastic ARIMA engine balances structural drift against auto-correlated moving dampeners. For this station, the projected median 10-year expectation remains bound to a steady baseline, treating past extreme shifts as stochastic anomaly shocks rather than permanent baseline acceleration."
  ))
})
```

